---
title: "Spotify Songs"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
source_code: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(tidyverse)
library(DT)
library(shiny)
library(ggplot2)
library(knitr)

dataPath <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-21/spotify_songs.csv'
loadData <- function(dataPath){
  if (!exists("data/spotify_songs.csv")){
    system("mkdir -p data")
    spotify_songs <- readr::read_csv(dataPath)
    write.csv(spotify_songs, "data/spotify_songs.csv")
  }else{
    spotify_songs <- read.csv("data/spotify_songs.csv")
  }
  return(spotify_songs)
}
spotify <- loadData(dataPath)
```
About
=====================================  
The purpose of this dashboard is to explore how features of songs in different genres vary. As well, we are interested in comparing the overall popularity of different genres of music and how the musical features affect popularity. The data and data dictionary can be found [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-01-21).

    
Column {data-width=600}
-------------------------------------

```{r}
library("vembedr")
link <- "https://youtu.be/AvPrD-rqTV0"
embed_url(link) %>%
  div(class = "vembedr")
```


The Data {data-orientation=rows}
=====================================     
Row {data-height=75}
-------------------------------------
This dataset consists of songs from Spotify playlists of specific genres, as well as musical features for each of the songs such as the rhythm, danceability, and loudness. This dataset also includes the popularity for each of the songs.

Row {data-height=600}
-------------------------------------
### Taking a look at the data
```{r}
tableData <- spotify %>%
  select(!c(track_id, track_album_release_date, playlist_id, track_album_id))
DT::renderDataTable(tableData)
```


Interactive: Musical features by genre {data-orientation=rows}
=====================================     
   
Row {data-height=600}
-------------------------------------
Column {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput("genre", label = "Select a genre:",
            choices = unique(spotify$playlist_genre), selected = "rock")

```
Column {data-height=1000}
-----------------------------------------------------------------------
```{r}
renderPlot({genreFeats <- spotify %>%
  select(danceability, energy, speechiness, acousticness, liveness, instrumentalness, playlist_genre) %>%
  filter(playlist_genre==input$genre)%>%
  select(danceability, energy, speechiness, acousticness, liveness, instrumentalness) %>%
  colMeans()%>%
  as.data.frame() %>%
  rownames_to_column("feature") %>%
  rename("avg"= ".")
  p <- ggplot(genreFeats, aes(x=feature, y=avg))+
  geom_bar(stat="identity")+
  xlab("Musical Feature")+
  ylab("Average value")
  
  print(p)})
```
Interactive: Random song generator {data-orientation=rows}
=====================================     
   
Row {data-height=600}
-------------------------------------
Column {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput("genre2", label = "Select a genre to get a random song from that genre:",
            choices = unique(spotify$playlist_genre), selected = "rock")

```
Column {data-height=1000}
-----------------------------------------------------------------------
```{r}
renderText({
  songNames <- spotify %>%
    filter(playlist_genre == input$genre2)%>%
    select(track_name, track_artist)
  ind <- sample(1:length(songNames$track_name), 1)
  name <- songNames[[ind,1]]
  artist <- songNames[[ind,2]]
  print(paste(name, "by", artist))

})

```


Static: Genre popularity {data-orientation=rows}
=====================================     
Column {.tabset}
-----------------------------------------------------------------------


### Popularity
```{r}
spotifyPop <- spotify %>%
  group_by(playlist_genre) %>%
  select(track_popularity, playlist_genre)%>%
  group_modify(~ as.data.frame(mean(.$track_popularity)))

colnames(spotifyPop) <- c("genre", "averagePop")

renderPlot({ggplot(spotifyPop, aes(x=genre, y=averagePop))+
  geom_bar(stat="identity")+
  xlab("Genre")+
  ylab("Average popularity")+
  ggtitle("Average popularity of songs in each genre")})
```


### Playlist

```{r, out.extra=c('allow="encrypted-media"', 'allowtransparency="true"', 'frameBorder="0"')}
knitr::include_url("https://open.spotify.com/embed/playlist/5uPUkCs6k2TvYc2RXjJjYs?utm_source=generator", height = "380")
```

Static: Danceability  {data-orientation=rows}
=====================================     
```{r}
genres <- unique(spotify$playlist_genre)
filt <- spotify %>%
   select(track_name, danceability, track_popularity, playlist_genre)

p <- ggplot(as.data.frame(filt), aes(x=playlist_genre, y=danceability))+
    geom_boxplot()+
    xlab("Genre")+
    ylab("Danceability")

renderPlot({print(p)})
  
  
```

   


Analysis {data-orientation=rows}
=====================================     
   
Row {data-height=600}
-------------------------------------
<!-- We can see that this dataset consists of 6 different musical genres. Let's see how many entries there are for each genre: -->
```{r}
library(tidyverse)
spotifyGenres <- spotify %>%
  group_by(playlist_genre)%>%
  group_modify(~ as.data.frame(as.character(dim(.x)[[1]])))
colnames(spotifyGenres) <- c("Genre", "Frequency")
renderPlot({grid.table(spotifyGenres)})
```

<!-- It looks like the dataset is relatively balanced, with edm taking the top spot at 6043 entries. (I will not be sharing my thoughts on edm at this time.) -->
<!-- Now, let's look at the relationship between popularity and danceability across all genres: -->
```{r}
library(ggplot2)
library(gridExtra)
myGrobs <- c()
genres <- unique(spotify$playlist_genre)
for (genre in genres){
  filt <- spotify %>%
    filter(playlist_genre == genre) %>%
    select(track_name, danceability, track_popularity, playlist_genre)
  p1 <- ggplot(as.data.frame(filt), aes(x=danceability, y=track_popularity))+
    geom_point()+
    xlab("Danceability")+
    ylab("Popularity")+
    ggtitle(stringr::str_to_title(genre))+
    geom_smooth()
  myGrobs[[genre]] <- p1
  
}
renderPlot({purrr::invoke("grid.arrange", c(myGrobs, ncol=3))})
```

<!-- *Figure 1:* Scatter plot of each track's danceability against the track's popularity, split by genre. -->


<!-- It seems that across all genres except edm, the popularity of a song increases as the danceability increases.This is surprising because I thought edm listeners really enjoy dancing. -->

<!-- Next, let's just take a quick look of the average of all features across the different genres: -->
Row {data-height=600}
-------------------------------------
```{r}
# spotifyFeats <- spotify %>%
#   select(track_name, danceability, energy, loudness, speechiness, acousticness, tempo, duration_ms, liveness, instrumentalness, playlist_genre)%>%
#   group_by(playlist_genre)
sp <- c()
for (genre in genres){
  sp[[genre]] <- spotify %>%
  select(danceability, energy, speechiness, acousticness, liveness, instrumentalness, playlist_genre) %>%
    filter(playlist_genre==genre)%>%
    select(danceability, energy, speechiness, acousticness, liveness, instrumentalness)
}
spMeans <-map(sp, colMeans)
spMeans <- map(spMeans, as.data.frame)
spMeans <- do.call(cbind, spMeans)
colnames(spMeans) <- genres
spMeans$features <- rownames(spMeans)
spMeansPlot <- spMeans %>%
  pivot_longer(cols=genres, names_to = "genre", values_to="averageValue")
renderPlot({ggplot(spMeansPlot, aes(x=features, y=averageValue))+
  geom_bar(stat="identity")+
  facet_wrap(~ stringr::str_to_title(genre))+
  theme(axis.text.x=element_text(angle = 90))+
  ylab("Average Value")+
  ggtitle("Average value of features across genres")})
```

<!-- *Figure 2*: Bar plot of average feature value for each genre. -->


<!-- Now let's look at the average popularity of each genre: -->
```{r}
spotifyPop <- spotify %>%
  group_by(playlist_genre) %>%
  select(track_popularity, playlist_genre)%>%
  group_modify(~ as.data.frame(mean(.$track_popularity)))
colnames(spotifyPop) <- c("genre", "averagePop")
renderPlot({ggplot(spotifyPop, aes(x=genre, y=averagePop))+
  geom_bar(stat="identity")+
  xlab("Genre")+
  ylab("Average popularity")+
  ggtitle("Average popularity of songs in each genre")})
```

<!-- *Figure 3:* Bar plot of the average popularity for each genre. -->


<!-- Unsurprisingly, pop is the most popular genre, and edm is the least popular on average.  -->

<!-- Overall, it seems that regardless of genre, listeners really enjoy danceable music. As well, across genres, it seems like the features of the music do not differ much, although Rap does have a higher average speechiness value than other genres. -->

